### 2014-01-28 conclusion.tex
2013.01.23 (Vassilis) Text refinements in introduction and conclusion sections.

% TODO: Should be reviewed
\section{Conclusion}
\label{section:conclusion}

Research and development around application systems that follow the 
\p\ architectural paradigm, have flourished in the last ten years or so.
By and large, such systems exploit the loose-coupling and self-organization 
of participating nodes, to shape substrates upon which diverse applications 
can run; the latter offer services including, but not limited to, 
distributed naming, 
Internet telephony and video conferencing, 
storage and indexing, 
content and file sharing, 
web crawling and caching, 
event notification and subscribing/publishing.
These services are provided atop the 
best-effort infrastructure of today's Internet 
and they have to effectively deal with an array of key quality attributes that
they may need or have to offer, such as node availability, robustness,
scalability, load-sharing, quality-of-service and user anonymity.
The difficulty to handle such features effectively, emanates from the fact that
peer nodes operate at the edge of the Internet and no adequate attention to the
structure of the physical network is paid during application-level network
formation. The deviation of the structure of the overlay network from the
optimal, according to the underlying physical structure, constitutes what is
termed topology mismatch and the problem is known as the topology mismatch
problem.

Development of \p--systems and applications
typically involve juggling a set of tradeoffs (e.g.,
choosing higher accuracy at the cost of increased system overhead.).
% Computer systems design is typically about juggling a set of tradeoffs (e.g.,
% choosing higher accuracy at the cost of increased system overhead.)  
In surveying more than
a decade's worth of research efforts aimed at solving the topology mismatch
problem in both unstructured and structured \p\ networks, we find this
tussle-of-tradeoffs property to hold.  
With regards to the three criteria we used --efficiency,
overhead, and scalability-- we find that none of the proposed solutions 
is superior to the others on all three fronts.  This is not surprising.
Nonetheless, by
\begin{inparaenum}[\itshape i\upshape)]
  \item presenting an analysis of each of the solutions,
including what distinguishes it from the
others and its advantages and disadvantages and
  \item offering a pictorial
comparison of how each solution fares compared with the others with regards
to efficiency, overhead, and scalability,
\end{inparaenum}
we hope to have provided \p\
developers and researchers with enough perspective so that depending
on the particular problem in hand, they are able to make a
correct choice amongst these surveyed here or, by firmly standing on
the shoulders of giants, propose a new solution of their own.