### 2014-01-28 background.tex
2013.01.24 (Vassilis) Addressing Alex's comments in previous iteration for background and conclusion sections.

\section{Background}
\label{section:background}

%The abstraction of the overlay network has been proposed as a way to implement
%efficient, fully distributed, application-layer services such as  
%routing messages to destinations that are not known in advance or building
%one-to-many addressing schemes 
% TODO find citation for Narada & Adobe's Real Time Media Flow Protocol (RTMFP)
%\cite{} that provide quality-of-service (QoS) guarantees. \emph{IP
%multicast}, as well as other proposals such as \emph{IntServ} and
%\emph{DiffServ} \cite{cisco_diffserv_2005}, have still not seen wide deployment
%largely due to the fact they require support from the IP
%infrastructure. On the other hand, an overlay network can be deployed on
%end-systems running the overlay protocol software, without the cooperation of ISPs.
%Academic research includes
%\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

In this section, we provide some background information on the types of
\p\ architectures that have been implemented and deployed. We also
describe and motivate more formally the topology mismatch problem, 
whose various aspects have been tackled over the past several years.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (\p) Overlay Architectures}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% VM: Don't know if we should start with this Gnutell specific material maybe
% it needs to go to the respective paragraph of dece.unstructured bellow.
Circa 2000, pioneering file-sharing \p-system,
Gnutella~\cite{gnutellav04,gnutella} introduced the \emph{servent }concept, a
\emph{portmanteau} that blends the notion of server
and client to denote the dual role of a node participating in its network.
Such server-less systems have proven to be able to achieve outstanding
aggregate resource capacities, as more and more participants join the system,
but without the additional expenditure for infrastructure.
While certain undesirable features such as 
free-riding~\cite{saroiu_measurefileshare_2002,adar_gnutellafreeriders_2000,hughes_gnutellafreeride_2005},
the distribution of illegal content and other~\emph{socio-technical}
issues \cite{hughes_socp2p_2008}, have arisen, \p\ systems have
continued to gain popularity throughout the years, primarily 
because they have successfully supported popular applications 
(i.e. file-sharing) amongst \emph{vast} numbers of participating users.
%%
For over a decade, excitement over the immense potential of \p\ systems has
generated a flurry of research and development, resulting in a wide range of
popular protocols, networks, and applications. As \p\ systems evolved, three
main architectures have emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

Designers of \emph{centralized} architectures 
(Figure~\ref{figure:p2p-archs:centralized}) were the first 
to recognize that requests for resources 
(e.g., CPU cycles or popular content) need not be fulfilled by
a dedicated server. Instead, requests could be satisfied by
many hosts that already possessed the resources in question. 
The downside of this approach was that it required a
centralized search service, that was instantly rendered a single point of
failure,  a scalability bottleneck, and a target of malicious acts including
\emph{DoS attacks}.
% \emph{Denial-of-Service (DoS) attacks}).

{\sl Napster} was the most successful incarnation of the
centralized approach, in file-sharing~\cite{napster}\footnote{Ofcourse, the
SETI@home project is considered the pioneer, but on a different application
domain \cite{seti}.}.
It maintained a \emph{central index server} based on file-lists 
that were peer-provided.
Users would query the index server that would ultimately 
yield pointers to the actual content.
Thus, by centralizing search, while distributing downloads, {\sl Napster}
achieved a highly functional design that, at its height, attracted $26$ million
users, sharing approximately $80$ million music
files\cite{jmm_naptopusage_2001}.
%%
The central indexing, though, played a vital role in the demise of 
{\sl Napter}. After \emph{Recording
Industry Association of America} successfully argued in court
copyright infringement of the artists' work, it was then easy to point to the
indexing server and ultimately bring it down.
%
% Napter's  centralized nature was what ultimately brought it down after Recording
% Industry Association of America (RIAA) filed a lawsuit for copyright infringement
% of the intellectual property of artists whose music was distributed by the
% Napster application. RIAA was able to point to the centralized indexing
% component of Napster as the entity responsible for illegal song-trading.

Architectures that distributed both search and resource provision 
capabilities (Figure~\ref{figure:p2p-archs:unstructred-full}) soon followed.
In these \emph{decentralized} approaches, 
resource placement within the overlay topology is random~\cite{YG-M2002};
for this reason, such architectures are referred to as \emph{unstructured}. 
Key properties of such \p\ systems are that they
support inherent heterogeneity of peers, 
are highly resilient to peer failures,
and incur low maintenance overhead at handling 
the dynamics of peer participation \cite{stutzbach_churn_2006}. 
Unstructured systems are also known as 
\emph{broadcast-based} systems as they use message flooding among
peers to propagate search queries. 
In its first realization,
{\sl Gnutella}(\emph{v.0.4}) is a well-known example of the fully
decentralized unstructured approach.

A key disadvantage of the unstructured approach is that message flooding burdens
the network since queries travel within the network randomly, visiting nodes
that do not have the sought resource and thus wasting bandwidth.  
To save on bandwidth, unstructured networks typically limit how long 
a query travels within the overlay via the \emph{time-to-live} mechanism. 
When a query is received by a node, it decreases its \emph{TTL} parameter by one
unit. If the \emph{TTL} value has not reached zero, the node forwards the
message to its neighbors; otherwise, it drops it.
While this approach prevents queries from visiting the entire network, it
does not guarantee that the resource of interest, (e.g., a rare recording) 
% of a musical group from the 1930s) 
will, ultimately, be found.
%%%
\begin{figure}[ht]
\centering
\subfigure[Centralized ({\sl SETI@Home} or {\sl Napster}). Peers
query the centralized 
entity (dotted lines) and then establish the connection to the data source.] {
  \includegraphics[scale=0.20]{img/pdf/centralized.pdf}
  \label{figure:p2p-archs:centralized}
}\qquad\qquad
\subfigure[Fully unstructured ({\sl Gnutella} \emph{v.0.4}). Data is distributed randomly,
the topology is also random and all participants are equal.] {
  \includegraphics[scale=0.20]{img/pdf/unstructured-full.pdf}
  \label{figure:p2p-archs:unstructred-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured ({\sl Gnutella} \emph{v.0.6} or {\sl FastTrack}). Certain
peers (bigger circles) shoulder specific responsibilities (e.g., data
indexing).] {
  \includegraphics[scale=0.20]{img/pdf/unstructured-hierarchical.pdf}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad\qquad
\subfigure[Structured ({\sl Chord} and {\sl Kademlia}). 
Peers and data are placed in the overlay in a deterministic manner.] {
  \includegraphics[scale=0.20]{img/pdf/structured.pdf}
  \label{figure:p2p-archs:structured}
}
\caption{The various \p\ architectures.}
\label{figure:p2p-archs}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To shorten search times, improve scalability, and lighten network load, 
unstructured configurations evolved into \emph{hierarchical} systems 
that dynamically assigned indexing functions to special peers, 
called \emph{ultra-}  or \emph{super-peers} 
(Figure~\ref{figure:p2p-archs:unstructured-hierarchical}). 
Peers with reliable network connections and/or high compute power 
became \emph{ultra/super-peers}
facilitating less-powerful \emph{leaf-nodes} to find resources of
interest. 
Examples of such hierarchical unstructured systems include
{\sl Gnutella}(\emph{v.0.6})/{\sl Limewire}~\cite{gnutella}, 
{\sl KaZaA}~\cite{kazaa} and {\sl Skype}~\cite{skype}. 
Although the hierarchical approach relieves network pressure as
search queries were flooded over the much smaller subset of 
ultra/super-peers, the disadvantages of reduced search scope and 
inability to locate rare resources remained.

The \emph{decentralized structured schemes}
(Figure~\ref{figure:p2p-archs:structured}) sought to address the 
disadvantages of the, above mentioned, decentralized unstructured
\p--approaches. Here, the main objective was to offer 
a self-organizing infrastructure for large-scale \p\ applications
\cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002,rgrk_bamboo_2004}.
\emph{Decentralized structured schemes}
implement a \emph{Distributed Hash Table (DHT)} that maps objects 
to nodes through a deterministic mechanism. 
\emph{DHT}s provide a guaranteed bound on the
number of overlay routing hops that have to be ``traveled''
% crossed 
to locate a resource, even in the case when only a \emph{single} copy exists.  
This requires
$O$($log(n)$) hops, compared to unstructured networks that require
$O(n)$ to reliably locate any object. 
Unfortunately this efficiency does not come without cost. 
Structured overlays can only support exact-match queries 
(i.e., queries that identify resources by name) as opposed
to content-based retrieval provided by unstructured overlays. 
This means that unstructured overlays can support more versatile 
resource location queries such as keyword and partial matching searches. 
Several popular file-sharing
\p\ applications have been based on \emph{DHT}s including
{\sl Overnet}/{\sl eDonkey}~\cite{overnet}, 
{\sl Kademlia}/{\sl Kad Network}~\cite{maymounkov_kademlia_2002}, and
some flavors of {\sl BitTorrent}~\cite{c_bittorrent_2003}.  
Moreover, \emph{DHT}s sparked
a flurry of research for many diverse of applications, ranging
from distributed search engines to event-notification and cloud-based
platforms~\cite{kbc_oceanstore_2000,rkcd_scribe_2001,mgpj_cloudsnap_2011}.

%
% TODO: SOME DISCUSSION ON P2P ARCHITECTURES
%
%\cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003}
%showed that the ad-hoc network topology of unstructured overlay networks that
%preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power
%Law describes the node degree while the Small World describes characteristics
%of path length and clustering coefficient. The clustering coefficient for a
%node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of
%the existing connections between $\upsilon$'s neighboring nodes to $\gamma
%\times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighboring
%nodes of $\upsilon$. High cluster coefficient means that neighboring nodes of
%any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999,
%saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured
%systems can have high maintenance traffic in delivering messages, updating the
%mapping, discovering failures and replicating lost data or pointers, making
%them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs
%there is a tight control between the data placement and the topology of the
%network. For this reason it is hard to efficiently support partially matched
%queries while Gnutella and other similar systems effortlessly support keyword
%searches and other complex queries since the mechanism is realized locally, on
%a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely
%for a flooding-based search to return results. DHTs, on the other hand, fit
%better in the systems which require ability to reliably locate content, even in
%the extreme case that only a single-copy exists in the network.
%\end{itemize}

%
% TODO: SOME DISCUSSION ON THE ADVANTAGES OF UNSTRUCTURED NETWORKS FROM
%       \cite{z-yk_ddno_2005}
%
%Unstructured P2P networks o?er a number of
%important advantages: (i) An unstructured network
%imposes very small demands on individual nodes,
%and more specifically?cally it allows nodes to join or leave
%the network without signi?cantly a?ecting the sys-
%tem performance. (ii) Unstructured networks are
%appropriate for content-based retrieval (e.g., key-
%word searches) as opposed to object identi?er loca-
%tion of structured overlays. (iii) Finally unstructured
%networks can easily accommodate nodes of vary-
%ing power. Consequently, they scale to very large
%sizes and they o?er more robust performance in
%the presence of node failures and connection
%unreliability.
%

%
% TODO: SOME DISCUSSION ON STRUCTURED OVERLAYS PROBABLY FROM
%       \cite{WLH2007}
% However, the design of a decentralized but structured P2P network has to
% overcome two critical issues. The first issue is the long routing latency.
% Several proximity schemes  have been proposed to avert long routing latency in
% current structured P2P networks, but they require a high-complexity procedure
% to periodically maintain the routing table (e.g. Pastry system) or they need
% pre-chosen landmarks to construct the overlay. However, the P2P system is by
% its very nature unstable since nodes join and leave frequently. For instance,
% the study of Gnutella shows around approximately 1200 membership changes per
% minute in a 100 000 nodes P2P system. Another proximity scheme needs some
% pre-chosen landmarks or a complete BGP routing table support. As a result,
% they both increase the difficulty of the P2P system deployment. The second
% issue is system maintenance overhead. The existing structured P2P networks
% allow nodes to keep some nearby nodes in their routing tables in order to
% achieve efficient routing.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Topology Mismatch Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

One of the major issues that defines the efficiency of 
an overlay network is its mapping to 
the underlying physical infrastructure. 
Recall, in Figure~\ref{figure:overlay} nodes $A$ and $B$ are 
on the same local network, while $C$ and $D$ are in different networks. 
The top layer represents the overlay interconnection formed by 
these four nodes at a higher level. 
Link connections there, can change as needed by 
the running environment, without any
particular constraint by the underlying physical topology. 
%%
For instance, assume nodes $A$, $B$, $C$ and $D$ being 
connected via the physical network shown in Figure~\ref{figure:phys}
in which costs are depicted in milliseconds. 
%%
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.4]{img/pdf/example-physical.pdf}
\caption{Interconnection of nodes in the physical level.}
\label{figure:phys}
\end{figure}
%%
When peer $A$ sends a message to its $D$ counterpart
and depending on the overlay configuration, 
nodes will experience different performance.
For example, in the overlay depicted in
Figure~\ref{figure:overlay-1}, the
message will traverse the following sequence of links in the physical layer
(marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2
\rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5
\rightarrow 5 \rightarrow 4$. In the alternative overlay of
Figure~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow
4$. 
The respective costs are $45 ms$ and $15 ms$.
% The total cost of the first sequence is $45 ms$, while the second yields a
% mere $15 ms$. 
Evidently, the second overlay is more
congruent with the underlying physical network in comparison 
to the first and, thus, is more efficient.

\begin{figure}[ht]
\centering
\subfigure[Nodes $B$ and $C$ have direct overlay connection.] {
  \includegraphics[scale=0.4]{img/pdf/example-overlay-inefficient.pdf}
  \label{figure:overlay-1}
}\qquad\qquad
\subfigure[Nodes $A$ and $D$ have direct overlay connection.] {
  \includegraphics[scale=0.4]{img/pdf/example-overlay-efficient.pdf}
  \label{figure:overlay-2}
}
\caption{Two different overlay connection configuration.}
\label{figure:overlay-confs}
\end{figure}

%% VM The next part was previously at the end of this paragraph and didn't seem
%% to fit right there. I moved it here since it is a theoretical optimality
%% definition used by some authors we review and I think it can help the reader
%% better understand what problem. Also I have rewritten it.
%%
The ratio of the physical distance (e.g., in hops) a query travels to
reach its target via overlay routing, to the actual minimal network distance via
IP routing, is termed in the literature as \emph{stretch} or \emph{Relative
Delay Penalty (RDP)}~\cite{CRZ2000}.
%%
Ideally, an overlay should be constructed to achieve an \emph{optimal} 
mapping between the overlay paths and the underlying network links to avoid
inefficient states where redundant physical resources are utilized for the
operation of the same link in the overlay.
The problem of constructing an optimal
overlay is referred to as the \emph{topology mismatch problem}, and is formally
defined as follows:
\begin{definition}
Let $V$=$\{v_1$, $\dots$, $v_n\}$ be a set of points denoting the network nodes,
$\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and
$v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch
problem is to construct a minimal spanning tree, where node degree is
restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Early incarnations of the overlay protocols did not make use of the optimal
mapping with the underlying network topology. {\sl Gnutella}, for example, is
considered far from scalable~\cite{ritter_gnucantscale_2001} since every peer
chooses its neighbors in the overlay without any knowledge of the underlying
network, resulting in mismatch. 
Queries may be flooded over multiple paths merging in one peer 
while one of the paths would suffice.  
Moreover, peers may forward the same message to each other
before they receive the query messages from each other.
\cite{matei_mapgnutella_2002} showed that, even when $95$\% of any two nodes 
are less than $7$ hops away from each other, 
a \emph{flooding-based} {\sl Gnutella}-like algorithm 
can generate $330$TB/month in a $50,000$ node network. 
A topology mismatch can thus have a serious effect 
on Internet router performance.

Similar problems are observed in decentralized structured schemes as well.
Typically, node IDs are assigned \emph{randomly}, resulting in excellent load
balancing, scalability, and robustness for the overlay. Unfortunately, this
randomness has a negative impact on the \emph{routing locality} of the network.
This means that even though the target node can be reached with a logarithmic
number of overlay hops, the distance traveled in the underlying network, 
during the overlay routing process, can be far from optimal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation and Goal for this Survey}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A topology--unaware overlay network is able to control the sequence of peers a
message traverses before reaching its destination, but it is completely unaware
of how the actual packets are switched at the underlying infrastructure along
the overlay path. 
In particular, a single logical point-to-point link on the
overlay, typically corresponds to multiple physical links in the
underlying layer. 
Moreover, a link in the physical network often lies
on several overlay paths inadvertently increasing the traffic on the
link; this is known as the link's \emph{stress}~\cite{CRSZ2002}. 
There is also the stochastic phenomenon during which peers 
depart and others arrive in the network, known 
as \emph{churn}~\cite{stutzbach_churn_2006}.
Even when an approach offers the best possible location
when a node joins in, the churn is set to  
continuously undermine the effort to maintain the best possible position
of the  peer in the network.

%% VM: This section has a lot of information that I belieave are worth
%% mentioning on the paper's conclusion section. We say for example why no
%% approach we reviewed is best of all (cause it is an NP-Hard problem); or why
%% virtual coordinate approaches will never be a reliable tool in today's
%% Internet. Also it doesn't add much to the purpose of this section that is to
%% say what motivated us to write the paper. If you also agree we can move it.
The topology mismatch problem is exacerbated by the extremely large amounts of
traffic generated by highly popular \p\ applications~\cite{byl-book}.
Addressing this problem remains a fundamental challenge 
in contributing to the overall efficiency of the Internet. 
Unfortunately, topology mismatch is known to be 
an \emph{NP-Hard} problem \cite{C2000,NPBOOK}. 
There is also additional complexity due to the fact that 
end-to-end latencies demonstrate \emph{triangle inequality
violations (TIVs)} which are a consequence of the Internet's structure and
routing policies; as such, they will remain a property of the Internet 
for the foreseeable future \cite{zheng_irprtt_2005}. 
\emph{TIV}s affect both network coordinate
\cite{cox_vivaldi_2004,wong_meridian_2005} and positioning \cite{ng_gnp_2001}
systems and makes the construction of \emph{IP}-topology-aware 
overlays difficult.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Over the past several years, researchers have investigated an extensive
array of heuristic solutions to the topology mismatch problem.
%% VM: If we move the above though we have to mention why all solutions are
%% heuristic.
The purpose of this survey is to gather and organize 
this extensive body of produced knowledge. 
We aspire to assist the reader in better understanding the 
different approaches suggested and to accomplish this, we 
present methods for both structured and unstructured distributed 
overlay networks.
We outline the respective algorithmic aspects and we categorize
the surveyed material so that pros and cons of each method
can become apparent in their context.

We employ both tabular and pictorial means to 
articulate features of the proposed approaches and outline 
both strengths and weaknesses.
To ascertain the effectiveness of each proposal, we predominantly use 
three criteria that capture the behavior of the algorithms 
discussed:
\emph{efficiency}, \emph{overhead} and \emph{scalability}.
For each of these criteria, we use a coarse-grained,
three-level, gradation to denote how well the algorithms do,
namely \emph{low}, \emph{medium} and \emph{high}.
Each criterion has its own
interpretation of the above literal values. 
\emph{Efficiency} means how well the
algorithm performs its search and topology adaptation functions (if any). There
can be many criteria that effect this metric, either positivelly or negatively.
For example, efficiency, is negatively effected every time an algorithm
reduces the system's average query search scope, or when the method employed
uses an unreliable metric to estimate physical proximity.
For the \emph{overhead} criterion, we roughly assess the cost
to compute and communicate peer join, message forwarding, 
and topology maintenance. As an example, algorithms that dictate only local
exchange of routing information during topology adaptation, can be safelly
considered to incur far less overhead, to the system, than those which
require global knowledge.
With \emph{scalability} we consider system behavior
with respect to the number of participating peers. We consider systems with low,
medium, and high scalability to mean they can drive hundreds, tenths of
thousands, and millions of nodes, respectively.
The heuristic nature of the solutions, inherently involve balancing trade-offs;
so we pay extra attention how enhancing the above criteria may, in response,
effect other aspects of the application. If, for example, tuning
topology-awareness of the BitTorrent environment results in the degradation of
the unrivaled optimal download performance of the vanilla-protocol, then this
cannot be considered a successful answer to the problem.

% we dedicate a discussion subsection after the presentation of the algorithms
% for both unstructured and structured architectures. This section offers:
% \begin{inparaenum}[\itshape i\upshape)]
  % \item further discussion that can help the reader identify, through
% a taxonomy, the pluses and minuses  of the presented algorithms,
  % \item a comprehensive tabular organization of the key-point information, and
  % \item an attempt to give a pictorial comparison of the relative performance of
% the algorithms.
% \end{inparaenum}

% We hope to offer the reader a deeper insight for the observations that
% pushed research toward some direction. Thus, we try to categorize these
% approaches so that we can organize directions making them comprehensible
% and helping the reader build critical thinking on the advantages and
% disadvantages of each approach.

% Tables provide a one-stop-shop overview of the algorithms and serve as a 
% tool both as a companion to the reader while reading the survey and a reference for
% future use.
% 
% Pictorial comparison consists of bar charts that give a relative feeling of the
% performance of the various algorithms, discussed in the corresponding section,
% in
% 
% terms of \emph{efficiency}, \emph{overhead} and \emph{scalability}; three
% important characteristics for every distributed algorithm. 
% For each of these criteria, we use a coarse-grained, 
% three-level, gradation to denote how well the algorithms do,
%%  namely \emph{low}, \emph{medium} and \emph{high}. 
% Each criterion has its own
% interpretation of the above literal values. Efficiency means how well the
% algorithm performs its search and topology adaptation functions, without
% affecting other characteristics of the networks (e.g. BitTorrent's optimal
% download performance). We assume that efficiency is
% negatively effected when an algorithm reduces the search scope. 
% %Moreover, the

%minimum-spanning-tree-based techniques have slow convergence speed in topology

%adaptation or the high clustering impacts load balancing and constrains search

%scope (or at least increases search satisfaction time), all reducing overall

%efficiency.