### 2015-07-26 cover-letter/comst/mard-coverletter-comst-ieee.tex
2015.07.23 (Alex) Renaming the paper to add the word "survey" in order to better fit JSS. Some minor fixes to author names and affiliations. Adding letter to JSS Editor-In-Chief and submission receipt + other resources.

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[12pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

% moderncv themes
\moderncvstyle{classic}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{blue}                              % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
%\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}                       % if you are not using xelatex ou lualatex, replace by the encoding you are using
%\usepackage{CJKutf8}                              % if you need to use CJK to typeset your resume in Chinese, Japanese or Korean

% adjust the page margins
\usepackage[scale=0.75]{geometry}
%\setlength{\hintscolumnwidth}{3cm}                % if you want to change the width of the column with the dates
%\setlength{\makecvtitlenamewidth}{10cm}           % for the 'classic' style, if you want to force the width allocated to your name and avoid line breaks. be careful though, the length is normally calculated to avoid any overlap with your personal info; use this at your own typographical risks...

% personal data
\name{Vassilis}{Moustakas}
\title{Software Development Engineer, M.Sc.}                               % optional, remove / comment the line if not wanted
\address{SDE at Amazon.com}{Marcel-Breuer-Str. 12}{80807 Munich, Germany}% optional, remove / comment the line if not wanted; the "postcode city" and and "country" arguments can be omitted or provided empty
\phone[mobile]{+49 (0) 176 55194610}                   % optional, remove / comment the line if not wanted
\phone[fixed]{+49 (0) 89 20097833}                    % optional, remove / comment the line if not wanted
%\phone[fax]{+3~(456)~789~012}                      % optional, remove / comment the line if not wanted
\email{vsmoustakas@gmail.com}                      % optional, remove / comment the line if not wanted
\homepage{de.linkedin.com/in/vsmoustakas}          % optional, remove / comment the line if not wanted
%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file
%\quote{Some quote}                                 % optional, remove / comment the line if not wanted

% to show numerical labels in the bibliography (default is to show no labels); only useful if you make citations in your resume
%\makeatletter
%\renewcommand*{\bibliographyitemlabel}{\@biblabel{\arabic{enumiv}}}
%\makeatother
%\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}% CONSIDER REPLACING THE ABOVE BY THIS

% bibliography with mutiple entries
%\usepackage{multibib}
%\newcites{book,misc}{{Books},{Others}}
%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
\recipient{Ekram Hossain}{Professor/Editor-In-Chief IEEE COMST\\Department of Electrical and Computer Engineering\\University of Manitoba\\75A Chancellor's Circle\\Winnipeg, MB R3T 5V6 CANADA}
\date{June 28, 2015}
\opening{Dear Professor Hossain,}
\closing{Sincerely,}
%\enclosure[Attached]{curriculum vit\ae{}}          % use an optional argument to use a string other than "Enclosure", or redefine \enclname
\makelettertitle

We are submitting a survey paper entitled ``Alleviating the Topology Mismatch Problem in Distributed Overlay Networks'' for publication in the Journal IEEE Communications Surveys \& Tutorials (IEEE COMST). The paper is co-authored with Assoc. Prof. H\"useyin Akcan of Izmir University of Economics in Izmir, Turkey, Prof. Mema Roussopoulos of the University of Athens in Athens, Greece and Prof. Alex Delis of the University of Athens in Athens, Greece.

Thank you very much for your kind consideration and your effort to have our work refereed in a timely fashion. We look forward to hearing from you.

\makeletterclosing

\end{document}

%% end of file `template.tex'.