### 2015-09-13 acknowledge.tex
2015.09.20 (Vassilis) Added acknowledgements section. Updated metric definitions. Introduced timeline figures to capture termoral and trending dimensions of surveyed material. Minor fixes and typos.

\section{Acknowledgements}
\label{section:acknowledge}

We would also like to express our gratitude to all the anonymous reviewers who helped us improve the presentation of our work with their insightful comments. This research has been partially supported by \emph{EC FP7 Workprogramme (2007-2013)} under grant agreement n. 283610 \emph{``European Scalable EarthScience Service Environment (EarthServer)''} and \emph{EC H2020} under grant agreement n. 654367 \emph{``Agile Analytics on Big Data Cubes (EarthServer 2)''}.

This  work  has  been  partially  supported  by \emph{SocWeb}, \emph{iMarine}, and \emph{Sucre} \emph{FP7 EU projects}.