### 2015-09-13 mard-survey.tex

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%% SP 2008/03/01

%%\documentclass[preprint,12pt,authoryear]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
\documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% packages

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

\usepackage{subfigure}
\usepackage{longtable}
\usepackage{adjustbox}
\usepackage{url}

\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{wasysym}
\usepackage{pdflscape}

\usepackage{chronosys}

\usepackage{pgfplots}
\usetikzlibrary{patterns}
\usepackage{supertabular}

\usepackage{comment}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Systems and Software}

\include{header}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Alleviating the Topology Mismatch Problem in Distributed Overlay Networks: A Survey}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

% \author{Vassilis Moustakas,
        % H\"useyin Akcan,
        % Mema Roussopoulos,
        % and~Alex Delis}
% 
% \address{
% Amazon Services\emph{vsmoustakas@gmail.com,}
% Izmir Economics Univ.  \emph{huseyin.akcan@ieu.edu.tr,}
% Univ. of Athens \emph{\{mema,ad\}@di.uoa.gr}
% }

\author[rvt0]{Vassilis Moustakas} % \fnref{fn1}}
\ead{vsmoustakas@gmail.com}
\author[focal]{H\"useyin Akcan} % \fnref{fn2}}
\ead{huseyin.akcan@ieu.edu.tr}
\author[rvt]{Mema Roussopoulos} % \fnref{fn3}}
\ead{mema@di.uoa.gr}
\author[rvt]{Alex Delis\corref{cor1}} % \fnref{fn3}}
\ead{ad@di.uoa.gr}
\cortext[cor1]{Corresponding author}
% \cortext[cor2]{Principal corresponding author}
% \fntext[fn1]{Department of Informatics and Telecommunications, University of Athens, Athens, 15703, Greece.}
% \fntext[fn2]{Department of Software Engineering, Izmir University of Economics, 35330 Balcova, Izmir, Turkey}
% \fntext[fn3]{Yet another author footnote. Indeed, you can have
% any number of author footnotes.}
\address[rvt0]{Amazon Development Center Germany GmbH, Marcel-Breuer Str. 12, 80807 Munich, Germany}
\address[focal]{Department of Software Engineering, Izmir University of Economics, 35330 Balcova, Izmir, Turkey}
\address[rvt]{Department of Informatics and Telecommunications, University of Athens, Athens, 15703, Greece}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
\input{abstract}
\end{abstract}

\begin{keyword}
Distributed systems, peer-to-peer, overlay network, topology mismatch,topology awareness

%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%% \linenumbers

%% main text

\input{intro}
\input{background}

\input{unstructured}

\input{structured}

\input{conclusion}

\input{acknowledge}

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% If you have bibdatabase file and want bibtex to generate the
%% bibitems, please use
%%
\bibliographystyle{elsarticle-harv} 
\bibliography{mard-survey}

%% else use the following coding to input the bibitems directly in the
%% TeX file.

%%\begin{thebibliography}{00}

%% \bibitem[Author(year)]{label}
%% Text of bibliographic item

%%\bibitem[ ()]{}

%%\end{thebibliography}
\end{document}

\endinput
%%
%% End of file `elsarticle-template-harv.tex'.