### 2015-09-26 journals/jss/bios/bios.tex

\documentclass[10pt,letterpaper]{article}

\usepackage{color}
\usepackage[margin=3cm, a4paper]{geometry}
\usepackage[hyphens]{url}
\usepackage{paralist}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

\begin{document}

\vspace*{0.5cm}
\centerline{\large \sf Biographical Sketches}
\vspace*{0.5cm}
\vspace*{0.5cm}

\emph{Vassilis Moustakas} received his MS Computer Science from the Univ. of Athens and his BS in Informatics from the Technological Educational Institute of Athens.  He is currently a member of the technical staff at the Amazon Development Center Germany in Munich as part of the A9.com Product Search group implementing scalable search and discovery technologies for Amazon.com and other websites. Prior to that, he worked as a software engineer for a number of companies as well as a researcher for various FP7 EU projects. His interests lie in Information Retrieval, Data Mining and Distributed Computing.

\emph{H\"useyin Akcan} is an Associate Professor in the Department of Software Engineering at Izmir University of Economics. His research interest are in computer systems, sensor networks, computational geometry and computer gaming. He has received both an PhD and an MS in Computer Science from the NYU Polytechnic School of Engineering and his BS from Instabul Technical University.

\emph{Mema Roussopoulos}  is an Associate Professor of Computer Science at the Univ. of Athens. Her interests  are  in  the  areas  of  distributed systems, networking, and digital preservation. She is a recipient of the NSF CAREER Award, the ERC Starting Grant Award, and Best Paper Award at ACM SOSP 2003.  She received her PhD in Computer Science from Stanford University and her BS from the Univ. of Maryland at College Park.

\emph{Alex Delis} is a Professor of Computer Science at the University of Athens. His research interests are in distributed computing and data management.  His work has been supported by agencies and organizations in Australia, US and the European Union. He holds both a PhD and an MS in Computer Science from the Univ. of Maryland at College Park and received his Eng. Diploma from the Univ. of Patras.

\end{document}