### 2015-09-26 journals/jss/highlights/highlights.tex

\documentclass[10pt,letterpaper]{article}

\usepackage{color}
\usepackage[margin=3cm, a4paper]{geometry}
\usepackage[hyphens]{url}
\usepackage{paralist}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

\begin{document}

\vspace*{0.5cm}
\centerline{\large\sf Highlights for JSS-D-15-00624}
\vspace*{0.5cm}
\vspace*{0.5cm}

\begin{itemize}
\item
The topology mismatch problem of P2P architectures is systematically defined.
\item
We aggregate research tackling the topology mismatch problem in P2P systems.
\item
Both unstructured and structured overlay paradigms are surveyed.
\item
We provide tabular data of salient features of all surveyed approaches.
% \item
% Abstracted metrics that help provide a pictorial comparison of methods.
\item
Trending and temporal characteristics of algorithms are presented.
% \item
% No definite solution found, rather a heuristic tussle-of-trade-offs property holds.
% \item
% A solid reference for system architects and researchers in the field is furnished.
\end{itemize}

\end{document}