### 2015-09-26 journals/jss/letter-to-editor/letter-to-editor-r1.tex

\documentclass[11pt]{article}

%% inclusion of packages
\usepackage{diletter-wsig}
% \usepackage{diletter-w-sig}
\usepackage{times}
\usepackage{url}

%% paragraph setting
\setlength{\parindent}{0pt}
\setlength{\parskip}{15pt}

%% style-specific parameters
\todayis{September 28th, 2015}
\myname{Alex Delis \\ Professor}
\myemail{\url{ad@di.uoa.gr}}
\myweb{\url{www.di.uoa.gr/~ad}}
\myphone{+30-210-727.5212}
\myfax{+30-210-727.5214}
\mycountry{Greece}
\mytitle{}

\begin{document}
\recipient{%
Professor Eleni Karatza \\
Area Editor - \emph{Journal of Systems and Software} \\
Department of Informatics \\
Aristotle University of Thessaloniki \\
54124 Thessaloniki, Greece
}

\begin{letter}{Dear Professor Karatza, }

\baselineskip=1.13\baselineskip

We are submitting the revision of our paper entitled 
``Alleviating the Topology Mismatch Problem in Distributed Overlay Networks
: A Survey''. We have initially submitted this paper to 
{\it Journal of Systems and Software} on July 21st, 2015.

The paper is co-authored with Vassilis Moustakas 
of Amazon Development Center in Munich, Germany, 
Huseyin Akcan of Izmir University of Economics in 
Izmir, Turkey,  and 
Mema Roussopoulos of the Univ. of Athens in Athens, Greece.

We also provide specific replies to comments made by the 
two reviewers.

Thank you very much for kindly taking the time to handle our 
revised submission and we look forward to hearing from you.

\myclosing{Sincerely yours,}

\end{letter}

\end{document}