### 2015-09-26 journals/jss/letter-to-editor/letter-to-editor.tex

\documentclass[11pt]{article}

%% inclusion of packages
\usepackage{diletter-wsig}
% \usepackage{diletter-w-sig}
\usepackage{times}
\usepackage{url}

%% paragraph setting
\setlength{\parindent}{0pt}
\setlength{\parskip}{15pt}

%% style-specific parameters
\todayis{July 21st, 2015}
\myname{Alex Delis \\ Professor}
\myemail{\url{ad@di.uoa.gr}}
\myweb{\url{www.di.uoa.gr/~ad}}
\myphone{+30-210-727.5212}
\myfax{+30-210-727.5214}
\mycountry{Greece}
\mytitle{}

\begin{document}
\recipient{%
Professor Hans van Vliet \\
Editor-in-Chief, \emph{Journal of Systems and Software}\\
Dept. of Computer Science, VU University, \\
De Boelelaan 1081, 1081 HV, \\
Amsterdam, The Netherlands
}

\begin{letter}{Dear Professor van Vliet,}

\baselineskip=1.13\baselineskip

We are submitting a paper entitled 
``Alleviating the Topology Mismatch Problem in Distributed Overlay Networks
: A Survey''
for consideration 
to the {\it Journal of Systems and Software}.

The paper is co-authored with Vassilis Moustakas 
of Amazon Development Center in Munich, Germany, 
Huseyin Akcan of Izmir University of Economics in 
Izmir, Turkey,  and 
Mema Roussopoulos of the Univ. of Athens in Athens, Greece.

Thank you very much for kindly taking the time to handle our submission and
we look forward to hearing from you soon.

\myclosing{Sincerely yours,}

\end{letter}

\end{document}